---
title: "SNAPSwitch assay"
author: "Haiyin Liu"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(openxlsx)
library(RColorBrewer)
library(patchwork)
library(ggrepel)
```

### Overview

Analysis of SNAPSwitchflow cytometry assay performed on the Cytek Aurora, recording AF488 (LNP association), Cy5 (SNAPSwitch probe or Cy5), and mScarlet (reporter expression) in live cells (DAPI negative).

### Files

#### Data file

`data.csv` is directly exported from Flowjo with these columns: AF488/Cy5/mScarlet gMFI, GROUPNAME (name starts with a number between 1-4, this is the `group_ID`), and TUBENAME (name is made up of `sample_ID` and replicate suffix `_001`, for example `LNP1_001`).

#### Sample key files

The script looks for key files in a `parameters` folder. It should contain

-   `groups.csv` takes the `group_ID` (1 to 4) and assigns the cell type (`WT` or `LSA`), and sensor (`SNAP` or `Cy5`).
-   `sample_info.csv` links `sample_ID` to the LNP name (e.g. batch, targeting) and formulation.
-   `SNAP_batches.csv` links the batch of SNAP oligo used with the experimentally determined correction factor, to account for switch-on efficacy.
-   `column_info.csv` should be updated with the actual column names in your `Table.csv`, and what they correspond to. You can assign them different colours in there too 

#### configuration files

Inside the `parameters` folder is a `config` folder for some more information and customisation.

-   `lipid_properties.csv` contains informations for ionisable lipids (e.g. pKa, lipid family) for lipid comparisons (e.g. 2D plots). *Note: lipid family is not used in calculations yet*
-   `R_colours.csv` assigns colour to R colour names used in different plot types

```{r parameters, message=FALSE}
#-------- YOUR EXPERIMENT PARAMETERS HERE --------#

datafile <- "20241210_Table.csv"                   # file name of your MFI table exported from FlowJo
exp_title <- "Thuy_test"        # date and name of the experiment
SNAP_batch <- 240527                      # ID of the SNAP-oligo batch that was used
formulation_order <- c("DMG"
                       )        # order in which formulations will be plotted if grouping is desired
untreated_ID <- "PBS"                   # untreated sample names for background subtraction
lipids <- FALSE                            # are you comparing lipids?

# load group and sample key files
facs_unclean <- read_csv(datafile)
group_info <- read_csv("parameters/groups.csv")
sample_info <- read_csv("parameters/samples.csv")
column_info <- read_csv("parameters/column_info.csv")
#-------------------------------------------------#  
```

```{r files, echo=FALSE, message=FALSE}
# load general configuration files
R_colours <- read_csv("parameters/config/R_colours.csv")
if (lipids == TRUE) {lipid_properties <- read_csv("parameters/config/lipid_properties.csv")}

# helper vector to order sample_names their sample_ID (eg "LNP1", "LNP2" etc)
sample_order <- factor(sample_info$sample_name, levels = sample_info$sample_name[order(sample_info$sample_ID)])

# fetch correction factor from the SNAP batch used
SNAP_batches <- read_csv("parameters/SNAP_batches.csv")
conversion_factor <- SNAP_batches$corr_factor[SNAP_batches$batch == SNAP_batch]

# create output directory
if (!dir.exists("output/")){dir.create("output/")}
if (!dir.exists(paste0("output/",exp_title))){dir.create(paste0("output/",exp_title))}

```

### Analysis

This code performs data cleaning, assigns sample names and cell types, and then performs following calculations: 

#### Raw values
-   LNP association (AF488 gMFI)
-   Protein expression (mScarlet gMFI)
-   gMFIs for each sample are compared with the "untreated cells" sample (= background), and if significantly different, background is subtracted. If not significant, `0` is assigned.
-   Group 2 (SNAP-samples LSA cells) data is reported in the main plots/files, all other values are plotted under QC plots

#### Calculated values
-   Endosomal escape
    -   for SNAP samples, SNAP gMFI are normalised to AF488 on WT cells for each LNP (formulation + Ab combination)     -   after normalisation,  each "LSA" sample is compared with the corresponding "WT" sample (= background), and 
    if significantly different, "WT" is subtracted. If not significant, `0` is assigned. Following values are reported:
      -   "absolute" escape (`adj_SNAP_abs`, WT subtracted SNAP signal)
      -   "relative" escape (`adj_SNAP_rel`, SNAP signal divided by AF488 signal)
      -   % escape efficiency (`SNAP_perc_of_Cy5`, relative SNAP divided by relative Cy5) 
-   mScarlet expression relative to
    -   LNP association
    -   absolute escape

```{r cleaning, echo=FALSE,message=FALSE}
####--------DATA CLEANING--------####
# trim data and rename columns
facs <- facs_unclean %>%
  select(-1) %>%
  select(where(function(x) !all(is.na(x)))) %>%
  filter(!grepl('mystery|Unstained|LNP10', TUBENAME)) %>% # removes "mystery" and Aurora "Unstained" tubes
  filter(!TUBENAME == "-")                                          # removes mean and SD rows

# rename columns and append "_raw" for fluorescent channels that need calculation
# use matching vector for renaming, then assign a column_ID for matching names
rename_vector <- setNames(column_info$column_ID, column_info$column_name)
colnames(facs) <- colnames(facs) %>%
  { ifelse(. %in% names(rename_vector), rename_vector[.], .) } %>%
  # append "_raw" where if_subtract is TRUE
  { ifelse(. %in% column_info$column_ID[column_info$if_subtract], paste0(., "_raw"), .) }

# generate group_ID column from the first number of GROUPNAME
facs <- add_column(facs,"group_ID" = as.numeric(sub(" .*", "", facs$GROUPNAME)))

# fetch and combine group_ID information to extract `cells`, `sensor` and `formulation` columns
facs <- facs %>% 
  left_join(group_info, by = "group_ID") %>%
  mutate("sample_ID" = word(TUBENAME,1,sep = "_")) %>% # extract sample_ID information from TUBENAME
  left_join(sample_info, by = "sample_ID") %>%
  # set the order in which the samples appear on the QC graphs
  mutate(sample_name = factor(sample_name, levels = sample_order))
```

```{r analysis, echo=FALSE, message=FALSE}
####--------CALCULATION--------####
# get fluorescence values for untreated samples
facs_untreated <- facs %>%
  filter(sample_ID == untreated_ID) %>%
  select(cells,AF488_raw,Cy5_raw,mScarlet_raw)

# function that tests for each cell type, treatment group, and sample, if values are significantly different from the untreated samples that cell type
calc_significant <- function(cell_type) {
  facs %>%
    filter(cells == cell_type) %>%
    group_by(group_ID, sample_ID) %>%
    mutate(
      significant_AF488 = t.test(AF488_raw, facs_untreated$AF488_raw[facs_untreated$cells == cell_type])$p.value < 0.05,
      significant_Cy5 = t.test(Cy5_raw, facs_untreated$Cy5_raw[facs_untreated$cells == cell_type])$p.value < 0.05,
      significant_mScarlet = t.test(mScarlet_raw, facs_untreated$mScarlet_raw[facs_untreated$cells == cell_type])$p.value < 0.05
    ) %>%
    ungroup()
}

# combine results for WT and LSA cells
facs_significant <- bind_rows(calc_significant("WT"), calc_significant("LSA"))

# if AF488 and mScarlet values are significantly different from background, subtract background (untreated_ID) to obtain true "signal"
# otherwise, set signal to 0 if not significantly different
facs_raw <- facs_significant %>%
  select(-c(GROUPNAME,TUBENAME)) %>%
  group_by(cells) %>%     # separate calculations for WT or LSA cells
  mutate(
    AF488 = if_else(significant_AF488, AF488_raw - mean(AF488_raw[sample_ID == untreated_ID]), 0),
    # CHANGE 251007: Cy5 is not reported as pure value, only LSA - WT, therefore subtraction only afterwards 
    # CHANGE 250321: Added back Cy5 subtraction for the Cy5-LNP ONLY 
    Cy5 = if_else(
      sensor == "Cy5",
      if_else(significant_Cy5, Cy5_raw - mean(Cy5_raw[sample_ID == untreated_ID]), 0),
      Cy5_raw
    ),
    mScarlet = if_else(significant_mScarlet, mScarlet_raw - mean(mScarlet_raw[sample_ID == untreated_ID]), 0)
    ) %>%
  ungroup() %>%
  # remove untreated_ID samples from dataframe
  filter(sample_ID != untreated_ID)

### code chunk to report samples where any fluorescence was found to be "not significant"
# false_results <- facs_raw %>%
#   filter(significant_AF488 == FALSE | significant_mScarlet == FALSE | significant_Cy5 == FALSE) %>%
#   select(sensor, cells, formulation, sample_name, significant_AF488, significant_mScarlet, significant_Cy5) %>%
#   rowwise() %>%  # Apply row-wise operations
#   mutate(
#     not_significant = paste(
#       c(
#         if (significant_AF488 == FALSE) "AF488" else NULL,
#         if (significant_mScarlet == FALSE) "mScarlet" else NULL,
#         if (significant_Cy5 == FALSE) "Cy5" else NULL
#       ),
#       collapse = ", "
#     )
#   ) %>%
#   ungroup() %>%
#   select(sample_name, not_significant, formulation, sensor, cells) %>%  # Reorder the columns
#   distinct() %>%  # Remove duplicate rows, if any
#   arrange(sample_name)  # Sort rows by sample_name
# 
# write.csv(false_results, "output/not_significant_samples.csv", row.names = FALSE) # Export summary

# calculations to normalise Cy5 to 488 MFI for both SNAP-LNP and Cy5-LNP.
facs_norm <- facs_raw %>%
  select(-c(AF488_raw,Cy5_raw,mScarlet_raw,significant_AF488,significant_Cy5,significant_mScarlet)) %>%
  group_by(sensor,formulation,sample_ID) %>%   # treat EACH formulation/sample combo as its own "formulation"
  mutate(norm_factor = AF488 / mean(AF488[cells == "WT"]),
         adj_Cy5 = Cy5 / norm_factor) %>% 
  ungroup()

# check if SNAP signal is significantly different in WT and LSA samples
facs_norm <- facs_norm %>%
  group_by(sample_ID) %>%
  mutate(
    significant_SNAP = ifelse(
      sensor == "SNAP",
      {
        WT_values <- adj_Cy5[cells == "WT" & sensor == "SNAP"]
        LSA_values <- adj_Cy5[cells == "LSA" & sensor == "SNAP"]
        
        # check for NA value, if any are NA or if there are fewer than 3 values, set significant_SNAP to FALSE
        if (anyNA(WT_values) || anyNA(LSA_values) || 
            length(WT_values) < 2 || length(LSA_values) < 2) {
          FALSE
        } else {
          t.test(WT_values, LSA_values)$p.value < 0.05
        }
      },
      NA
    )
  ) %>%
  ungroup()

# perform SNAP fluorescence calculation
facs_norm <- facs_norm %>%
  group_by(formulation, sensor, sample_ID) %>%
  mutate(

    # for SNAP samples, subtracts WT signal, don't subtract WT from Cy5)
    adj_Cy5_abs = ifelse(sensor == "SNAP", 
                         
                         # check if significant_SNAP is TRUE, then subtract WT from from LSA, otherwise set to 0
                         ifelse(significant_SNAP, 
                                adj_Cy5 - mean(adj_Cy5[cells == "WT"]), 
                                0), # Set to 0 if significant_SNAP is FALSE
                         
                         adj_Cy5), # Keep adj_Cy5 as is if sensor is Cy5
    adj_Cy5_abs = pmax(adj_Cy5_abs, 0), # Set negative values to 0
    adj_Cy5_rel = adj_Cy5_abs / AF488, # Divide by AF488 signal of the same sample
    express_per_assoc = mScarlet / AF488,
    express_per_escape = ifelse(adj_Cy5_abs == 0, NA, mScarlet / adj_Cy5_abs) # Set NA if adj_Cy5_abs is 0
  ) %>%
  ungroup()

# set negative escape samples to zero


# extract final results from the SNAP samples
results_SNAP <- facs_norm %>%
  # removes WT values as they have been subtracted from the respective LSA samples
  filter(cells != "WT") %>%
  # for each SNAP sample, calculate the % escape by dividing the relative SNAP/488 signal by the relative Cy5/488 signal  
  group_by(formulation,sample_name) %>%
  mutate(SNAP_perc_of_Cy5 = ifelse(sensor == "SNAP", adj_Cy5_rel[sensor == "SNAP"] * conversion_factor * 100 / mean(adj_Cy5_rel[sensor == "Cy5"]), NA)) %>%
  # drop Cy5 samples from final table
  filter(sensor == "SNAP") %>%
  ungroup()

# join results table with lipid properties
results_SNAP <- results_SNAP %>%
  {if (lipids == TRUE) left_join(., lipid_properties, by = c("sample_name" = "lipid_name")) else .}

```

### Output

This code creates exploratory graphs from the data and exports tables into a `results.xlsx` file in a prism-friendly format (for "Grouped" graphs) to copy paste if further adjustments (eg. combining replicate experiments) are needed. 

The variables to export are defined in `datarange`.

```{r variables}
####--------VARIABLES TO PLOT--------####
# define which columns should be used to generate plots and export
# adj_Cy5_rel is still calculated but not shown/exported as we now use SNAP_perc_of_Cy5
datarange <- c("AF488",
               "adj_Cy5_abs",
               "SNAP_perc_of_Cy5",
               "express_per_escape",
               "mScarlet")

datarange_raw <-c("AF488_raw", 
               "Cy5_raw", 
               "mScarlet_raw", 
               "viability")
```

```{r plots, echo=FALSE, message=FALSE, fig.width=9, fig.height=6}
####--------PLOTS--------####
# define colour of the plots for each variable
colourrange <- c("darkseagreen",
                 "skyblue2",
                 "grey68",
                 "mediumpurple3",
                 "indianred")

# TODO fix the colour range and make it match the datarange (figure out how to match it not only for the raw data channels) 
# column_colour <- column_info$colour
# colourrange <- data.frame(colour = column_colour) %>%
#  left_join(R_colours, by = "colour") %>%
#  arrange(match(colour, column_colour)) %>%
#  # if comparing only one "formulation" group, use a plain R colour, otherwise use the Brewer pallette
#  {if (length(formulation_order) == 1) {pull(., R_colour)} else {pull(., Brewer_colour)}}

# function to generate ggplots for each variable
bar_chart <- function(df,column,colour) {
  df %>% 
    select(c(formulation,sample_name,all_of(column))) %>%
    group_by(sample_name,formulation) %>%
    summarise(mean = mean(!!sym(column)),
              sd = sd(!!sym(column))) %>%
    ggplot(aes(
      x=if (lipids) reorder(sample_name, -mean) else factor(sample_name, levels = sort(unique(sample_name))),
      y=mean, 
      fill = factor(formulation, formulation_order))) +
    geom_bar(stat="identity", position = "dodge", fill = colour) +
    geom_errorbar(aes(ymin = mean - sd, ymax = mean + sd), 
                  position = position_dodge(0.9),
                  width = 0.3,
                  linewidth = 0.2,
                  colour = "#333333") +
    labs(x = NULL,
         y = column,
         fill = "formulation") +
    theme_minimal(base_size = 10) +
    theme(legend.key.size = unit(0.2, "cm"),
          axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1.1)
)
}

# generate list to populate with plots
plot_list <- sapply(datarange,function(x) NULL)

# generate plots for each variable in the datarange and match with the colour range
for(i in 1:length(datarange)){plot_list[[i]] <- bar_chart(results_SNAP, datarange[i], colourrange[i])}

# combine all plots in the list into an array (patchwork package)
all_plots <- wrap_plots(plot_list, ncol = 2) + 
  plot_annotation(title = exp_title, 
                  theme = theme(plot.title = element_text(size = 10)))

# save and display plots
ggsave(paste0("output/", exp_title,"/",exp_title,".png"), plot = all_plots, width = 9, height = 6, units = "in")
all_plots
```


### 2D plots

```{r 2D plots, echo=FALSE, fig.height=6, fig.width=9, message=FALSE}
if (lipids == TRUE) {
  colourrange_2D <- c("Greens",
                     "Greys",
                     "Purples",
                     "PuRd")

  results_2D <- results_SNAP %>%
    select(sample_name, pka, tails, unsat, all_of(datarange)) %>%
    filter(!pka == 99,
           !grepl('Mix', sample_name)) 
  
  ### PKA PLOTS
  # dot plots where the values are plotted against the pKa
  plots_pka <- function(df, column, colour) {
    df2 <- df %>% 
      select(c(sample_name, pka, all_of(column))) %>%
      group_by(sample_name, pka) %>%
      summarise(mean = mean(!!sym(column))) %>%
      ungroup()
    
    ggplot(df2, aes(x = pka, y = mean)) +
    geom_point(color = colour) +
    labs(x = "pKa",
         y = column) +
    geom_text_repel(
      aes(label = sample_name),
      size = 3,
      box.padding = unit(0.2, "lines"),
      point.padding = unit(0.2, "lines")
      ) +
    theme_minimal(base_size = 10) +
    theme(legend.key.size = unit(0.2, "cm"),
          panel.grid.major.y = element_blank(),  # Remove y-axis gridlines
          panel.grid.minor.y = element_blank(),
          axis.line = element_line(color = "black"))
  }
    
  # generate list to populate with plots
  plot_list_pka <- sapply(datarange,function(x) NULL)
  
  # generate plots for each variable in the datarange and match with the colour range
  for(i in 1:length(datarange)){plot_list_pka[[i]] <- plots_pka(results_2D, datarange[i], colourrange[i])}
  
  # combine all plots in the list into an array (patchwork package)
  all_plots_pka <- wrap_plots(plot_list_pka, ncol = 2) + 
    plot_annotation(title = paste0(exp_title, " - pKa plots"), 
                    theme = theme(plot.title = element_text(size = 10)))
  
  # save and display plots
  ggsave(paste0("output/", exp_title,"/",exp_title,"_pKa.png"), plot = all_plots_pka, width = 10, height = 8, units = "in")
  all_plots_pka
}

# these are obsolete now that we group them by family in prism
# TODO add the lipid families as another column in the lipid parameter file and group them them inside R
# ### 2D DOT PLOTS
# # dot plots with pKa vs tails and the value expressed as point size/colour
# plots_pkavtails <- function(df, column, colour) {
#   df2 <- df %>% 
#     select(c(sample_name, pka, tails, all_of(column))) %>%
#     group_by(sample_name, pka, tails) %>%
#     summarise(mean = mean(!!sym(column))) %>%
#     ungroup()
#   
#   # select top 3 points to label in the graph
#   top_points <- df2 %>% arrange(desc(mean)) %>% slice_head(n = 3)
# 
#   ggplot(df2, aes(y = tails, x = pka, color = mean)) +
#   geom_jitter(aes(size = mean),width = 0.05, height = 0) +
#   scale_color_distiller(palette = colour, trans = "reverse") +  # Reverse the color scale
#   scale_size_continuous(range = c(0.2, 3)) +
#   geom_text(data = top_points, aes(label = sample_name), vjust = -0.5, hjust = -0.1, size = 3) +
#   guides(color = guide_colorbar(order = 1, reverse = TRUE),
#          size = guide_legend(order = 2, title = NULL)) +
#   labs(x = "pKa",
#        y = "number of tails",
#        color = column) +
#   theme_minimal(base_size = 10) +
#   scale_y_continuous(limits = c(1.5, 4.5), breaks = c(2, 3, 4)) +
#   theme(legend.key.size = unit(0.2, "cm"),
#         panel.grid.major.y = element_blank(),  # Remove y-axis gridlines
#         panel.grid.minor.y = element_blank(),
#         axis.line = element_line(color = "black"))
# }
# 
# # generate list to populate with plots
# plot_list_2D <- sapply(datarange,function(x) NULL)
# 
# # generate plots for each variable in the datarange and match with the colour range
# for(i in 1:length(datarange)){plot_list_2D[[i]] <- plots_pkavtails(results_2D, datarange[i], colourrange_2D[i])}
# 
# # combine all plots in the list into an array (patchwork package)
# all_plots_2D <- wrap_plots(plot_list_2D, ncol = 2) + 
#   plot_annotation(title = exp_title, 
#                   theme = theme(plot.title = element_text(size = 10)))
# 
# # save and display plots
# ggsave(paste0("output/", exp_title,"_2D.png"), plot = all_plots_2D, width = 10, height = 8, units = "in")
# all_plots_2D

```



### QC plots

```{r QC plots, echo=FALSE, fig.height=6, fig.width=9, message=FALSE}
# extract raw data for QC and split them into SNAP and Cy5 samples 
results_raw <- facs_raw %>%
  select(c(
    "AF488_raw",
    "Cy5_raw",
    "mScarlet_raw",
    "viability",
    "formulation",
    "sample_name",
    "sample_ID",
    "cells",
    "sensor"
    )
  )
results_raw <- split(results_raw, results_raw$sensor)

# define colour of the plots for each variable
colourrange_raw <- c("darkseagreen",
                   "skyblue2",
                   "indianred",
                   "grey68")

## TODO find a way to include faceting for LSA and WT samples so the scales are the same
# function to generate ggplots for each QC variable using the raw data
QC_bar_chart <- function(df,column,colour) {
  df %>% 
    select(c(formulation,sample_name,cells,all_of(column))) %>%
    group_by(cells,sample_name,formulation) %>%
    summarise(mean = mean(!!sym(column)),
              sd = sd(!!sym(column))) %>%
    ggplot(aes(x=sample_name, y=mean, fill = factor(formulation, formulation_order))) +
    geom_bar(stat="identity", position = "dodge", fill = colour) +
    geom_errorbar(aes(ymin = mean - sd, ymax = mean + sd), 
                  position = position_dodge(0.9),
                  width = 0.3,
                  linewidth = 0.2,
                  colour = "#333333") +
    labs(x = NULL,
         y = column,
         fill = "formulation") +
    theme_minimal(base_size = 10) +
    theme(legend.key.size = unit(0.2, "cm"),
          axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1.1)
          ) +
    facet_wrap(vars(cells), scales = "free")
}

# generate list to populate with plots
SNAP_QC <- sapply(datarange_raw,function(x) NULL)
Cy5_QC <- sapply(datarange_raw,function(x) NULL)

# generate plots for each variable in the datarange and match with the colour range
for(i in 1:length(datarange_raw)){SNAP_QC[[i]] <- QC_bar_chart(results_raw[["SNAP"]], datarange_raw[i], colourrange_raw[i])}
for(i in 1:length(datarange_raw)){Cy5_QC[[i]] <- QC_bar_chart(results_raw[["Cy5"]], datarange_raw[i], colourrange_raw[i])}

# combine all plots in the list into an array (patchwork package)
SNAP_QC_plots <- wrap_plots(SNAP_QC[1:2], ncol = 1) + 
  plot_annotation(title = paste0(exp_title, " QC (SNAP)"), 
                  theme = theme(plot.title = element_text(size = 10)))
SNAP_QC_plots2 <- wrap_plots(SNAP_QC[3:4], ncol = 1)

Cy5_QC_plots <- wrap_plots(Cy5_QC[1:2], ncol = 1) + 
  plot_annotation(title = paste0(exp_title, " QC (Cy5)"), 
                  theme = theme(plot.title = element_text(size = 10)))
Cy5_QC_plots2 <- wrap_plots(SNAP_QC[3:4], ncol = 1)

SNAP_QC_plots
SNAP_QC_plots2
Cy5_QC_plots
Cy5_QC_plots2

# TODO think about how to best output the QC plots
```

```{r main tables, echo=FALSE, message=FALSE}
####--------EXCEL EXPORT--------####
# MAIN RESULT FILE
# function to organise the data in a prism-friendly format
prismformat <- function(df, column) {
  df %>%
  group_by(sample_name, formulation) %>%
  mutate(formulation = paste0(formulation,'_',row_number())) %>%
  ungroup() %>%
  select(c(formulation,sample_name,{{column}})) %>%
  pivot_wider(names_from = formulation, values_from = {{column}})
} 

# create empty list for main variables of interest ("datarange")
results_list <- sapply(datarange,function(x) NULL)
# populate list with results from the SNAP samples using the prismformat function
for(i in 1:length(datarange)){results_list[[i]] <- prismformat(results_SNAP, !!datarange[i])}

# create empty list for raw data
raw_list <- sapply(datarange,function(x) NULL)

# function to reorganise the data for copying into prism
prismrenumber <- function(df) {
  # assign numbers to the preferred order for the formulation
  mapping <- setNames(seq_along(formulation_order), formulation_order)
  for (pattern in names(mapping)) {
    colnames(df)[grep(pattern, colnames(df), ignore.case = TRUE)] <- 
      paste0(mapping[pattern], "_", colnames(df)[grep(pattern, colnames(df), ignore.case = TRUE)])
  }

    # reorder columns (with numbered formulation groups) and sort sample by name
  df <- df %>%
    select("sample_name", sort(setdiff(colnames(.), "sample_name"))) %>%
    arrange(sample_name)
  
  return(df)
}

# apply function to all dataframes in the list
results_list <- lapply(results_list, prismrenumber)

# determine lipid order sorted by best mScarlet expression
mScarlet_rank <- results_list$mScarlet %>%
  mutate(average = rowMeans(across(2:4)),
         rank = rank(-average, ties.method = "min")) %>%
  select(sample_name, rank)

# add rank to every dataframe in the list and save to excel
results_list <- lapply(results_list, function(df) {
  df %>%
    left_join(mScarlet_rank, by = "sample_name") %>%
    arrange(rank)  # Sort by rank in ascending order
})
write.xlsx(results_list, file = paste0("output/", exp_title,"/",exp_title,".xlsx"))

# COMBINED ANALYSIS FILE
# write a csv of the processed data to feed into the SNAPSwitch_combined analysis 
write.csv(results_SNAP, file = paste0("output/", exp_title, "/", exp_title,"_SNAP.csv"))

# PLOT FILES FOR PKA ETC
# create table with pKa and other characteristics of each sample vs the means of each variable in datarange
if (lipids == TRUE) {
  results_SNAP_pka <- results_SNAP %>% 
    group_by(sample_name,pka,tails,unsat) %>% 
    summarise(across(.cols = all_of(datarange), mean))
  write.csv(results_SNAP_pka, file = paste0("output/", exp_title,"/",exp_title,"_pKa.csv"))
}
```
